---
title: "DATA2020FinalProject_plan"
author: "Hanjun Wei, Keying Gong, Yurui Zhang"
date: "4/26/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data

The data our group uses is the Police Shootings in the US: <https://www.washingtonpost.com/graphics/investigations/police-shootings-database/>

### Question

Our target question: What will the total number of fatal shootings by on duty police officers all over the united states next month? 

### Our brief plan

The objective we are interested in:
1. Analyze historical fatal shootings based on demographic groups
2. predict 2022 fatal shootings using time-series model for the whole population and for different racial groups

#### Step 1: data cleaning and conduct feature engineering 
#### Step 2: Exploratory Data Analysis on the police shooting dataset. Discover interesting patterns
#### Step 3: Compare race composition of each state VS. fatal shootings race composition by state
#### Step 4:

(Time Series Modeling)
build ARIMA model to predict 2022 fatal shootings using time-series model for the whole population and for different racial groups


(Racial Mixed Effect Modeling)
Reconstruct the dataset in terms of Race. 


Feature #1 is Racial Category, which includes 7 types of race.


Feature #2 is Number of Fatal Shooting at each time stamp, count the total number of fatal shootings in a given time interval.


Feature #3 is Time Stamp, month and year.
Construct a mixed effect modeling in terms of the number of fatal shootings.


(State Mixed Effect Modeling)
Reconstruct the dataset in terms of State. 


Feature #1 is State Category, which includes 50 types of state.


Feature #2 is Number of Fatal Shooting at each time stamp, count the total number of fatal shootings in a given time interval.


Feature #3 is Time Stamp, month and year.
Construct a mixed effect modeling in terms of the number of fatal shootings.




```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load the needed packages
library(dplyr)
library(caret)
library(glmnet)
library(MASS)
library(InformationValue)
library(tidyverse)
library(naniar)
library(ggplot2)
library(lubridate)
library(visdat)
library(finalfit)
library(GGally)
library(ggmap)
library(tidygeocoder)
library(usmap)
library(reshape2)
```

### Data Cleaning first

First thing first, read the data into the file.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Load the needed dataset
data_shoot <- read_csv("~/Desktop/Brown/Brown_class/2022_Spring/DATA2020/Final_project/data-police-shootings-master/fatal-police-shootings-data.csv")

# make a copy of the readed in data just in case
data_shoot_not_change <- data_shoot
```

Print out the first 10 rows to better understand the data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
head(data_shoot, 10)
```

Looks good to me.

let us print the names of all of the columns to see if it matches our key table.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
names(data_shoot)
```

Then after we read in the data, we can take a look at some characteristic of the data, First of all, we print out the shape of the data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(dim(data_shoot))
```

We have 7291 rows and 17 columns.

we want to know the exact percentage of missing value in each columns, so we print out the percentage of not missing values in our datasets.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Missing data - to check the percent of not missing data points in each columns
apply(data_shoot, 2, function(x) sum(complete.cases(x))/nrow(data_shoot))
```

From the plot, we can see that there are missing values in name, armed, age, gender, race, flee, longitude, and latitude.

Then we will run the missing pattern plots to take a better look at the missineness.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
vis_miss(data_shoot)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
missing_pattern(data_shoot)
```

For our variable "id", from the code book, this should be the continuous variable that contains the id information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
length(unique(data_shoot$id))
```

For our variable "name", from the code book, this should be the categorical variable that contains the id information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable name
length(unique(data_shoot$name))
```

For our variable "date", from the code book, this should be the date variable. So we add the Year, Month and WeekDay new variables.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
length(unique(data_shoot$date))

data_shoot$Year <- year(data_shoot$date)
data_shoot$Month <- month(data_shoot$date)
data_shoot$WeekDay <- weekdays(data_shoot$date)

data_shoot$Year <- as.factor(data_shoot$Year)
data_shoot$Month <- as.factor(data_shoot$Month)
data_shoot$WeekDay <- as.factor(data_shoot$WeekDay)

```

For our variable "manner_of_death", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
unique(data_shoot$manner_of_death)
data_shoot$manner_of_death <- as.factor(data_shoot$manner_of_death)
```

For our variable "armed", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
# long_dis_death, long_dis_not_death, short_dis_death, short_dis_not_death,


long_dis_death <- c("gun","guns and explosives","crossbow","gun and knife","hatchet and gun","machete and gun","gun and sword","gun and car","pellet gun","bow and arrow","gun and vehicle","vehicle and gun","grenade","air pistol","gun and machete")
long_dis_not_death <- c("toy weapon","nail gun","BB gun","Taser","bean-bag gun","hand torch","pepper spray","fireworks","incendiary device","Airsoft pistol")
short_dis_death <- c("knife","hatchet","sword","machete","sharp object","meat cleaver","straight edge razor","ax","chain saw","scissors","pick-axe","spear","pitchfork","glass shard","metal rake","pole and knife","chainsaw","samurai sword","baseball bat and knife","ice pick","machete and hammer")
short_dis_not_death <- c("shovel","hammer","box cutter","metal object","screwdriver","lawn mower blade","flagpole","cordless drill","metal pole","metal pipe","metal hand tool","blunt object","metal stick","chain","contractor's level","railroad spikes","stapler","beer bottle","binoculars","baseball bat and fireplace poker","brick","baseball bat","garden tool","pipe","flashlight","baton","chair","rock","piece of wood","pole","crowbar","oar","tire iron","air conditioner","baseball bat and bottle","pen","wrench","walking stick","barstool","wasp spray","bottle","microphone","stake")
vehicles <- c("vehicle","carjack","motorcycle","BB gun and vehicle","vehicle and machete","car, knife and mace","knife and vehicle")
undetermined <- c("undetermined","unknown weapon","claimed to be armed")
unarmed <- c("unarmed")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% long_dis_death, "armed_level"] <- "long_dis_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% long_dis_not_death, "armed_level"] <- "long_dis_not_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% short_dis_death, "armed_level"] <- "short_dis_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% short_dis_not_death, "armed_level"] <- "short_dis_not_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% vehicles, "armed_level"] <- "vehicles"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% undetermined, "armed_level"] <- "undetermined"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% unarmed, "armed_level"] <- "unarmed"

data_shoot$armed_level <- as.factor(data_shoot$armed_level)

```

For our variable "age", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
# divide into several ranges



# 0-18 pre-young, 18-35 young adulthood 35-55 middle age 55 older adulthood 

unique(data_shoot$age)

data_shoot[is.na(data_shoot$age)== FALSE & data_shoot$age > 0 & data_shoot$age <= 18, "age_group"] <- "(0-18) pre-young"


data_shoot[is.na(data_shoot$age)== FALSE & data_shoot$age > 18 & data_shoot$age <= 35, "age_group"] <- "(18-35) young adulthood"


data_shoot[is.na(data_shoot$age) == FALSE & data_shoot$age > 35 & data_shoot$age <= 55, "age_group"] <- "(35-55) middle age"


data_shoot[is.na(data_shoot$age) == FALSE & data_shoot$age > 55, "age_group"] <- "(>55) older adulthood"


unique(data_shoot$age_group)

data_shoot$age_group <- as.factor(data_shoot$age_group)
```

For our variable "gender", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable gender
unique(data_shoot$gender)
data_shoot$gender <- as.factor(data_shoot$gender)
```

For our variable "race", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
unique(data_shoot$race)
data_shoot$race <- as.factor(data_shoot$race)
```

For our variable "city", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
length(unique(data_shoot$city))
```

For our variable "state", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
unique(data_shoot$state)
data_shoot$state <- as.factor(data_shoot$state)

north_east <- c("MA", "RI", "CT", "VT", "NH", "ME", "PA", "NJ", "NY")
south_east <- c("DC", "GA", "NC", "SC", "VA", "WV", "KY", "TN", "MS", "AL", "DE", "MD", "FL", "LA", "AR")
mid_west <- c("MN", "WI", "IL", "OH", "IN", "MI", "MO", "IA", "KS", "NE", "ND", "SD")
south_west <- c("NM", "AZ", "OK", "TX")
the_west <- c("CA", "CO", "NV", "HI", "AK", "OR", "UT", "ID", "MT", "WY", "WA")

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% north_east, "state_loc"] <- "north_east"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% south_east, "state_loc"] <- "south_east"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% mid_west, "state_loc"] <- "mid_west"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% south_west, "state_loc"] <- "south_west"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% the_west, "state_loc"] <- "the_west"

unique(data_shoot$state_loc)

data_shoot$state_loc <- as.factor(data_shoot$state_loc)

```

For our variable "signs_of_mental_illness", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$signs_of_mental_illness)
data_shoot$signs_of_mental_illness <- as.factor(data_shoot$signs_of_mental_illness)
```

For our variable "threat_level", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$threat_level)
data_shoot$threat_level <- as.factor(data_shoot$threat_level)
```

For our variable "flee", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$flee)
data_shoot$flee <- as.factor(data_shoot$flee)
```

For our variable "body_camera", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

unique(data_shoot$body_camera)
data_shoot$body_camera <- as.factor(data_shoot$body_camera)
```

Now we have to create two new datasets, one for eda and analysis, one for creating model.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
names(data_shoot)
```


### Bring some outside data source

US States by Race 2022 from world population review
Link: https://worldpopulationreview.com/states/states-by-race


```{r, echo=FALSE, message=FALSE, warning=FALSE}

state_race <- read_csv("~/Desktop/Brown/Brown_class/2022_Spring/DATA2020/Final_project/state_race.csv")

```

Let's reshape the data to better join the table

```{r, echo=FALSE, message=FALSE, warning=FALSE}
state_race_reform <- melt(state_race, id = c("State","Total"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
names(state_race_reform)<- c('state', 'Total_population_by_state', 'race', 'Race_population_by_state')

state_race_reform
```


```{r}
state_race_reform_a <- state_race_reform %>% dplyr::select(state, Total_population_by_state)
state_race_reform_b <- state_race_reform %>% dplyr::select(state, race, Race_population_by_state)

data_shoot <- merge(data_shoot, state_race_reform_b, by=c("state","race"),all.x=TRUE)


sum(is.na(data_shoot$Total_population_by_state))
```




US States personal income from bea.gov
Link: https://www.bea.gov/data/gdp/gdp-state


```{r, echo=FALSE, message=FALSE, warning=FALSE}

state_income <- read_csv("~/Desktop/Brown/Brown_class/2022_Spring/DATA2020/Final_project/Personal_income_by_states.csv")

```

Let's reshape the data to better join the table


```{r, echo=FALSE, message=FALSE, warning=FALSE}
state_income_reform <- melt(state_income, id = c("state"))

names(state_income_reform)<- c('state', 'year_qtr', 'personal_income_sum_whole_state')
state_income_reform
```



```{r}
library(zoo)
data_shoot$year_qtr <- format(as.yearqtr(data_shoot$date, format = "%Y-%m-%d"), format = "%Y:Q%q")

data_shoot <- merge(data_shoot, state_income_reform, by=c("state","year_qtr"),all.x=TRUE)

summary(data_shoot)

```

### EDA

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert to Date (for time series plot)
data_shoot_eda <- data_shoot
data_shoot_eda$count <- 1

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(summary(data_shoot_eda))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot, aes(x = manner_of_death)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = gender)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = race)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = signs_of_mental_illness)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = threat_level)) +
  geom_bar()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = body_camera)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = Year)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = armed_level)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = state_loc)) +
  geom_bar()
```



```{r}

data_shoot_eda
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

theme_set(theme_classic())
# Histogram on a Categorical variable
g <- ggplot(data_shoot_eda, aes(state_loc))
g +geom_bar(aes(fill=race), width = 0.5) + 
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9)) + 
  xlab("States Regions") +
    ylab("Fatal Shootings") +
  labs(title="Stacked Barplot on Categorical Variable", 
       subtitle="State Regions across Different Race")

```








```{r, echo=FALSE, message=FALSE, warning=FALSE}



ggplot(data_shoot_eda, aes(x=age, color=gender)) +
  geom_histogram(fill="white", position="dodge")+
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9),legend.position="top") + 
  xlab("Age") +
    ylab("Fatal Shootings") +
  labs(title="Histogram Plot on Categorical Variable", 
       subtitle="Age across Gender")



```




```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x=race, y= age)) + 
        geom_boxplot()

```

```{r}

ggplot(data = data_shoot_eda, aes(x = state_loc, y = count)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")



```

```{r}
data_shoot_eda

library(dplyr) 
data_shoot_eda <-data_shoot_eda %>% distinct()
```

```{r}

library(forcats)

# Creating Race table
Race_data_no_population <- data_shoot_eda %>% group_by(state) %>% summarise(counts = sum(count))

Race_data_no_population

Race_data_no_population <- Race_data_no_population[order(-Race_data_no_population$counts),]

Race_data_no_population %>%
  mutate(state = fct_reorder(state,counts)) %>%
  slice(1:10) %>%
  ggplot(aes(x=reorder(state,counts), y=counts)) +
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9),legend.position="top") +
    geom_bar(stat="identity",fill = "#4682B4", alpha=.6, width=.4) +
  geom_text(aes(label=counts), hjust=0, color="black", size=3.5)+
    coord_flip() +
    xlab("States") +
    ylab("Total Fatal Shootings") +
  labs(title="Barplot of Total Fatal Shootings", 
       subtitle="Top 10 States with Most Total Fatal Shootings")+
    theme_bw()

```







```{r}

# Creating Race table


names(state_race)<- c('state', 'Total_population_by_state', 'H', 'W', 'B','N','A',"O")

Race_data_with_population <- merge(Race_data_no_population, state_race, by=c("state"),all.x=TRUE)
Race_data_with_population$counts_per <- round(Race_data_with_population$count/Race_data_with_population$Total_population_by_state * 1000,digits = 5)



Race_data_with_population <- Race_data_with_population[order(-Race_data_with_population$counts_per),]

  
Race_data_with_population%>%
slice(1:10) %>%
  ggplot(aes(x=reorder(state,counts_per), y=counts_per)) +
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9),legend.position="top") +
    geom_bar(stat="identity",fill = "#EA9999", alpha=.6, width=.4) +
  geom_text(aes(label=counts_per), hjust=0, color="black", size=3.5)+
    coord_flip() +
    xlab("States") +
    ylab("Fatal shootings every 1000 people in the state") +
  labs(title="Barplot of Fatal Shootings every 1000 people from 2015-2021", 
       subtitle="Top 10 States with Most Total Fatal Shootings every 1000 people from 2015-2021")+
    theme_bw()


```


```{r}
plot_usmap(data = Race_data_no_population, values = "counts",labels=TRUE,label_color = "white") + 
scale_fill_viridis_c(name = "Total Fatal Shootings")+
  theme(legend.position = "right")+
  labs(title="Map plot of total Fatal Shootings", 
       subtitle="Total Fatal Shootings ")



```


```{r}

Race_data_with_population


plot_usmap(data = Race_data_with_population, values = "counts_per",labels=TRUE,label_color = "white") + 
scale_fill_viridis_c(name = "Fatal Shootings every 1000 people")+
  theme(legend.position = "right")+
  labs(title="Map plot of Fatal Shootings every 1000 people", 
       subtitle="Fatal Shootings every 1000 people")

```








```{r}

data_by_race <- data_shoot_eda %>% group_by(race) %>% summarise(counts = sum(count))

data_by_race %>%
  mutate(race = fct_reorder(race, counts)) %>%
  ggplot(aes(x=race, y=counts)) +
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9),legend.position="top") +
    geom_bar(stat="identity",fill = "#93c47d", alpha=.6, width=.4) +
  geom_text(aes(label=counts), hjust=0.2, color="black", size=3.5)+
    coord_flip() +
    xlab("Race") +
    ylab("Total fatal shootings") +
  labs(title="Barplot of Total Fatal Shootings", 
       subtitle="Total Fatal Shootings for different races")+
    theme_bw()


```
```{r}

data_by_race
```



```{r}
data_by_race$total_in_us <- c(17708954,39977554,58479370,2160378,9271166,197100373,10^100)
data_by_race$counts_per_1000 <- round(data_by_race$counts/data_by_race$total_in_us * 1000,digits=5)

data_by_race %>%
  mutate(race = fct_reorder(race, counts_per_1000)) %>%
  ggplot(aes(x=race, y=counts_per_1000)) +
  theme(axis.text.x = element_text(size = 12, angle=0, vjust=0.9),legend.position="top") +
    geom_bar(stat="identity",fill = "#8e7cc3", alpha=.6, width=.4) +
  geom_text(aes(label=counts_per_1000), hjust=0.2, color="black", size=3.5)+
    coord_flip() +
    xlab("Race") +
    ylab("Fatal shootings every 1000 people in such race") +
    labs(title="Barplot of Fatal Shootings every 1000 people from 2015-2021", 
       subtitle="Fatal Shootings every 1000 people for different races from 2015-2021")+
  
    theme_bw()

data_by_race
```



```{r}
data_state_loc <- data_shoot_eda  %>% group_by(state_loc,state) %>% summarise(counts = sum(count))

data_state_loc <- merge(data_state_loc, state_race, by=c("state"),all.x=TRUE)

state_income$mean_over_time <- rowMeans(state_income[,-1])

data_state_loc <- merge(data_state_loc, state_income, by=c("state"),all.x=TRUE)


data_state_loc$counts_per <- data_state_loc$counts/data_state_loc$Total_population_by_state * 1000
data_state_loc$mean_over_time_per <- data_state_loc$mean_over_time/data_state_loc$Total_population_by_state * 1000000


# Scatterplot
gg <- ggplot(data_state_loc, aes(x=mean_over_time_per, y=Total_population_by_state)) + 
  geom_point(aes(col=state_loc, size=counts_per)) + 
  xlim(c(30000, 100000)) + 
  ylim(c(0, 10000000)) + 
  labs(subtitle="Per Capita Mean Annual Personal income Vs Population", 
       y="Population of State", 
       x="Per Capita Mean Annual Personal income ($dollars)", 
       title="Scatterplot")

gg



```






```{r}
geo(address = c("Tokyo", "Lima", "Nairobi"),
 method = 'osm')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(rworldmap)
# get map
worldmap <- getMap(resolution = "coarse")
plot(worldmap, xlim = c(-80, 160), ylim = c(-50, 100), 
     asp = 1, bg = "lightblue", col = "black", fill = T)
points(data_shoot$Longitude, data_shoot$Latitude, 
       col = "red", cex = .01)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_loc <- ggplot(data_shoot_eda, aes(x=longitude, y=latitude)) + geom_point()
print(plot_loc + ggtitle("police killed people locations all over us"))

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Creating Race table
data_shoot_eda_race <- data_shoot_eda%>% group_by(state) %>% summarise(counts = sum(count)/Total_population)
data_shoot_eda_race
```





################################################################################## Code ############################################################################################

## Time series EDA for race and states
```{r}
## Time trend plot for race

# Convert to Date (for time series plot)
time_data = data_shoot
time_data$count = 1
time_data$date = as.Date(paste(format(time_data$date, "%Y-%m"),"-01",sep=""))

# Creating Race table
Race_data = time_data%>% group_by(date, race) %>% summarise(counts = sum(count))

# Plot shooting number over time interns of race
p = ggplot(Race_data, aes(x = date, y = counts)) + geom_line(aes(color = race), size = 1) 
p + geom_area(aes(color = race, fill = race),  alpha = 0.5, position = position_dodge(0.8)) + xlab("Time") + ylab("Number of Shooting") + ggtitle("Variation of Police Shooting Death between 2015 and 2022 (Race)")


# We see there is huge increase in undefined race shooting number (problematic), let's only consider cases happen between 2015 and 2020
Race_data_reduced = Race_data %>% filter(date < '2021-01-01')


# Plot shooting number over time interns of race
p = ggplot(Race_data_reduced, aes(x = date, y = counts)) + geom_line(aes(color = race), size = 1) 
p + geom_area(aes(color = race, fill = race),  alpha = 0.5, position = position_dodge(0.8)) + xlab("Time") + ylab("Number of Shooting") + ggtitle("Variation of Police Shooting Death between 2015 and 2020 (Race)")

# without AUC
p = ggplot(Race_data_reduced, aes(x = date, y = counts)) + geom_line(aes(color = race), size = 1) 
p  + xlab("Time") + ylab("Number of Shooting") + ggtitle("Variation of Police Shooting Death between 2015 and 2020 (Race)")
```




```{r}
## Time trend for Region

# Convert to Date
time_data = data_shoot
time_data$count = 1
time_data$date = as.Date(paste(format(time_data$date, "%Y-%m"),"-01",sep=""))

# Race table
time_data = time_data%>% group_by(date, state_loc) %>% summarise(counts = sum(count))

# Plot
p = ggplot(time_data, aes(x = date, y = counts)) + geom_line(aes(color = state_loc), size = 1) 
p + geom_area(aes(color = state_loc, fill = state_loc),  alpha = 0.5, position = position_dodge(0.8)) + xlab("Time") + ylab("Number of Shooting") + ggtitle("Variation of Police Shooting Death between 2015 and 2022 (Region)")

# Plot without AUC
p = ggplot(time_data, aes(x = date, y = counts)) + geom_line(aes(color = state_loc), size = 1) 
p+ xlab("Time") + ylab("Number of Shooting") + ggtitle("Variation of Police Shooting Death between 2015 and 2022 (Region)")
```



## Data Table

```{r}
## Recreate dataset For Time Series

# Convert to Date
# Because We are group data, we need to choose our variables wisely.
shoot_num = data_shoot
shoot_num$count = 1
shoot_num$date = as.Date(paste(format(shoot_num$date, "%Y-%m"),"-01",sep=""))


shoot_num = shoot_num %>% group_by(date, race, state_loc) %>% summarise(counts = sum(count))

# Assign ID
#shoot_num
id = 1
for (i in unique(shoot_num$race)){
  for (j in unique(shoot_num$state_loc)){
    shoot_num[shoot_num$race %in% i & shoot_num$state_loc %in% j,"ID"] = id
    id = id + 1

  }
}
shoot_num$ID = as.factor(shoot_num$ID)
shoot_num = data.frame(shoot_num %>% dplyr::select(ID, date, race, state_loc, counts))
head(shoot_num)
```
```{r}
# select observations between 2015 to 2021
shoot_num = shoot_num %>% filter(date <'2021-01-01')
shoot_num
```

