---
title: "DATA2020FinalProject_plan"
author: "Hanjun Wei, Keying Gong, Yurui Zhang"
date: "4/26/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data

The data our group uses is the Police Shootings in the US: <https://www.washingtonpost.com/graphics/investigations/police-shootings-database/>

### Question

The question we are interested in:
1. Analyze historical fatal shootings based on demographic groups
2. predict 2022 fatal shootings using time-series model for the whole population and for different racial groups

### Our brief plan



#### Step 1: data cleaning and conduct feature engineering 
#### Step 2: Exploratory Data Analysis on the police shooting dataset. Discover interesting patterns
#### Step 3: Compare race composition of each state VS. fatal shootings race composition by state
#### Step 4:

(Time Series Modeling)
build ARIMA model to predict 2022 fatal shootings using time-series model for the whole population and for different racial groups


(Racial Mixed Effect Modeling)
Reconstruct the dataset in terms of Race. 


Feature #1 is Racial Category, which includes 7 types of race.


Feature #2 is Number of Fatal Shooting at each time stamp, count the total number of fatal shootings in a given time interval.


Feature #3 is Time Stamp, month and year.
Construct a mixed effect modeling in terms of the number of fatal shootings.


(State Mixed Effect Modeling)
Reconstruct the dataset in terms of State. 


Feature #1 is State Category, which includes 50 types of state.


Feature #2 is Number of Fatal Shooting at each time stamp, count the total number of fatal shootings in a given time interval.


Feature #3 is Time Stamp, month and year.
Construct a mixed effect modeling in terms of the number of fatal shootings.







```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load the needed packages
library(dplyr)
library(caret)
library(glmnet)
library(MASS)
library(InformationValue)
library(tidyverse)
library(naniar)
library(ggplot2)
library(lubridate)
library(visdat)
library(finalfit)
library(GGally)
library(ggmap)
library(tidygeocoder)
library(usmap)
```

### Data Cleaning first

First thing first, read the data into the file.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Load the needed dataset
data_shoot <- read_csv("~/Desktop/Brown/Brown_class/2022_Spring/DATA2020/Final_project/data-police-shootings-master/fatal-police-shootings-data.csv")

# make a copy of the readed in data just in case
data_shoot_not_change <- data_shoot
```

Print out the first 10 rows to better understand the data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
head(data_shoot, 10)
```

Looks good to me.

let us print the names of all of the columns to see if it matches our key table.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
names(data_shoot)
```

Then after we read in the data, we can take a look at some characteristic of the data, First of all, we print out the shape of the data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(dim(data_shoot))
```

We have 7291 rows and 17 columns.

we want to know the exact percentage of missing value in each columns, so we print out the percentage of not missing values in our datasets.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Missing data - to check the percent of not missing data points in each columns
apply(data_shoot, 2, function(x) sum(complete.cases(x))/nrow(data_shoot))
```

From the plot, we can see that there are missing values in name, armed, age, gender, race, flee, longitude, and latitude.

Then we will run the missing pattern plots to take a better look at the missineness.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
vis_miss(data_shoot)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
missing_pattern(data_shoot)
```

For our variable "id", from the code book, this should be the continuous variable that contains the id information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
length(unique(data_shoot$id))
```

For our variable "name", from the code book, this should be the categorical variable that contains the id information.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable name
length(unique(data_shoot$name))
```

For our variable "date", from the code book, this should be the date variable. So we add the Year, Month and WeekDay new variables.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
length(unique(data_shoot$date))

data_shoot$Year <- year(data_shoot$date)
data_shoot$Month <- month(data_shoot$date)
data_shoot$WeekDay <- weekdays(data_shoot$date)

data_shoot$Year <- as.factor(data_shoot$Year)
data_shoot$Month <- as.factor(data_shoot$Month)
data_shoot$WeekDay <- as.factor(data_shoot$WeekDay)

```

For our variable "manner_of_death", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable manner_of_death
unique(data_shoot$manner_of_death)
data_shoot$manner_of_death <- as.factor(data_shoot$manner_of_death)
```

For our variable "armed", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
# long_dis_death, long_dis_not_death, short_dis_death, short_dis_not_death,


long_dis_death <- c("gun","guns and explosives","crossbow","gun and knife","hatchet and gun","machete and gun","gun and sword","gun and car","pellet gun","bow and arrow","gun and vehicle","vehicle and gun","grenade","air pistol","gun and machete")
long_dis_not_death <- c("toy weapon","nail gun","BB gun","Taser","bean-bag gun","hand torch","pepper spray","fireworks","incendiary device","Airsoft pistol")
short_dis_death <- c("knife","hatchet","sword","machete","sharp object","meat cleaver","straight edge razor","ax","chain saw","scissors","pick-axe","spear","pitchfork","glass shard","metal rake","pole and knife","chainsaw","samurai sword","baseball bat and knife","ice pick","machete and hammer")
short_dis_not_death <- c("shovel","hammer","box cutter","metal object","screwdriver","lawn mower blade","flagpole","cordless drill","metal pole","metal pipe","metal hand tool","blunt object","metal stick","chain","contractor's level","railroad spikes","stapler","beer bottle","binoculars","baseball bat and fireplace poker","brick","baseball bat","garden tool","pipe","flashlight","baton","chair","rock","piece of wood","pole","crowbar","oar","tire iron","air conditioner","baseball bat and bottle","pen","wrench","walking stick","barstool","wasp spray","bottle","microphone","stake")
vehicles <- c("vehicle","carjack","motorcycle","BB gun and vehicle","vehicle and machete","car, knife and mace","knife and vehicle")
undetermined <- c("undetermined","unknown weapon","claimed to be armed")
unarmed <- c("unarmed")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% long_dis_death, "armed_level"] <- "long_dis_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% long_dis_not_death, "armed_level"] <- "long_dis_not_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% short_dis_death, "armed_level"] <- "short_dis_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% short_dis_not_death, "armed_level"] <- "short_dis_not_death"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% vehicles, "armed_level"] <- "vehicles"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% undetermined, "armed_level"] <- "undetermined"

data_shoot[is.na(data_shoot$armed)==FALSE & data_shoot$armed %in% unarmed, "armed_level"] <- "unarmed"

data_shoot$armed_level <- as.factor(data_shoot$armed_level)

```

For our variable "age", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
# divide into several ranges

unique(data_shoot$age)

data_shoot[is.na(data_shoot$age)==FALSE & data_shoot$age > 14 & data_shoot$age <= 44, "age_group"] <- "15-44"
data_shoot[is.na(data_shoot$age) == FALSE & data_shoot$age > 44 & data_shoot$age <= 64, "age_group"] <- "45-64"
data_shoot[is.na(data_shoot$age) == FALSE & data_shoot$age > 64, "age_group"] <- "> 64"


unique(data_shoot$age_group)

data_shoot$age_group <- as.factor(data_shoot$age_group)
```

For our variable "gender", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable gender
unique(data_shoot$gender)
data_shoot$gender <- as.factor(data_shoot$gender)
```

For our variable "race", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
unique(data_shoot$race)
data_shoot$race <- as.factor(data_shoot$race)
```

For our variable "city", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
length(unique(data_shoot$city))
```

For our variable "state", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For our variable armed
unique(data_shoot$state)
data_shoot$state <- as.factor(data_shoot$state)

north_east <- c("MA", "RI", "CT", "VT", "NH", "ME", "PA", "NJ", "NY")
south_east <- c("DC", "GA", "NC", "SC", "VA", "WV", "KY", "TN", "MS", "AL", "DE", "MD", "FL", "LA", "AR")
mid_west <- c("MN", "WI", "IL", "OH", "IN", "MI", "MO", "IA", "KS", "NE", "ND", "SD")
south_west <- c("NM", "AZ", "OK", "TX")
the_west <- c("CA", "CO", "NV", "HI", "AK", "OR", "UT", "ID", "MT", "WY", "WA")

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% north_east, "state_loc"] <- "north_east"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% south_east, "state_loc"] <- "south_east"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% mid_west, "state_loc"] <- "mid_west"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% south_west, "state_loc"] <- "south_west"

data_shoot[is.na(data_shoot$state)==FALSE & data_shoot$state %in% the_west, "state_loc"] <- "the_west"

unique(data_shoot$state_loc)

data_shoot$state_loc <- as.factor(data_shoot$state_loc)

```

For our variable "signs_of_mental_illness", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$signs_of_mental_illness)
data_shoot$signs_of_mental_illness <- as.factor(data_shoot$signs_of_mental_illness)
```

For our variable "threat_level", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$threat_level)
data_shoot$threat_level <- as.factor(data_shoot$threat_level)
```

For our variable "flee", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
unique(data_shoot$flee)
data_shoot$flee <- as.factor(data_shoot$flee)
```

For our variable "body_camera", from the code book, this should be the categorical variable.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

unique(data_shoot$body_camera)
data_shoot$body_camera <- as.factor(data_shoot$body_camera)
```

Now we have to create two new datasets, one for eda and analysis, one for creating model.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
names(data_shoot)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_shoot_eda <- data_shoot %>% dplyr::select(c("manner_of_death","age","gender","race","city","state","signs_of_mental_illness","threat_level","flee","body_camera","longitude","latitude","is_geocoding_exact","Year","Month","WeekDay","armed_level","age_group","state_loc"))

data_shoot_model <- data_shoot %>% dplyr::select(c("date","manner_of_death","age","gender","race","city","state","signs_of_mental_illness","threat_level","flee","body_camera","longitude","latitude","is_geocoding_exact","armed_level","age_group","state_loc"))

```


### EDA So far

```{r, echo=FALSE}
print(summary(data_shoot_eda))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = manner_of_death)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = gender)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = race)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = signs_of_mental_illness)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = threat_level)) +
  geom_bar()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = body_camera)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = Year)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = armed_level)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x = state_loc)) +
  geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
hist(data_shoot_eda$age)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_shoot_eda, aes(x=race, y= age)) + 
        geom_boxplot()

```

```{r}
geo(address = c("Tokyo", "Lima", "Nairobi"),
 method = 'osm')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(rworldmap)
# get map
worldmap <- getMap(resolution = "coarse")
plot(worldmap, xlim = c(-80, 160), ylim = c(-50, 100), 
     asp = 1, bg = "lightblue", col = "black", fill = T)
points(data_shoot$Longitude, data_shoot$Latitude, 
       col = "red", cex = .01)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_loc <- ggplot(data_shoot_eda, aes(x=longitude, y=latitude)) + geom_point()
print(plot_loc + ggtitle("police killed people locations all over us"))

```






